---
title: "EDA Assignment1"
author: "Gary"
date: "11/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This assignment uses data from the UC Irvine Machine Learning Repository, a popular repository for machine learning datasets. In particular, we will be using the “Individual household electric power consumption Data Set” which I have made available on the course web site.

Measurements of electric power consumption in one household with a one-minute sampling rate over a period of almost 4 years. Different electrical quantities and some sub-metering values are available.


```{r echo = F, load required libraries}
require(tidyverse)
```

## Read in and clean data
```{r read in data}
d <- read.csv(file = 'household_power_consumption.txt', sep = ';')


#filter only dates that are between 2007-02-01 and 2007-02-02
d <- d %>% filter(Date %in% c('1/2/2007', '2/2/2007'))

str(d)
```

There seems to be some data hygene issue, that some columns should be numeric but
read in as Factor due to missing data that is marked as "?".

We will attempt to clean it by replacing "?" with NA value and transform 
those columns into numeric. 

Afterwards, we will transform appropriate time series data into date/time format.

### Cleaning data 
```{r pressure, echo=FALSE}
#replace where there are "?" with NA value
d[d =='?'] <- NA

#Transform column 3 to 9 into numeric.
d[,3:9] <- apply(d[,3:9], 2, as.numeric)

#now transform time series data into appropriate forms
d$date2 <- as.Date(d$Date, format = '%d/%m/%Y')
date_time <- paste0(as.character(d$Date), ' ', as.character(d$Time))
d$datetime <- strptime(date_time, '%d/%m/%Y %H:%M:%S')

str(d)
```

Now data is cleaner, we can start making graphs

#Plot1
```{r, plot1}
hist(d$Global_active_power, 
     col = 'red',
     xlab = 'Global Active Power (kilowatts)',
     main = 'Global Active Power' )


```

#Plot2
```{r, plot2}
plot(x = d$datetime, 
     y = d$Global_active_power, 
     type = 'l',
     xlab = '',
     ylab = 'Global Active Power (kilowatts)')
```


#Plot3
```{r, plot3}
plot(x = d$datetime, 
     y = d$Sub_metering_1, 
     type = 'l',
     xlab = '',
     ylab = 'Global Active Power (kilowatts)',
     col = 'black')
par(new = T)
lines(x = d$datetime, 
     y = d$Sub_metering_2, 
     col = 'red')
par(new = T)
lines(x = d$datetime, 
     y = d$Sub_metering_3, 
     col = 'blue')

legend( 'topright', 
       legend = c('Sub_meeting_1','Sub_meeting_2', 'Sub_meeting_3'),
       col = c('black', 'red', 'blue'),
       lty = 1, 
       cex = 1)
par(new = F)
```



#Plot4
```{r, plot4}
par(mfrow = c(2,2))

#plot 1.1
plot(x = d$datetime, 
     y = d$Global_active_power, 
     type = 'l',
     xlab = '',
     ylab = 'Global Active Power (kilowatts)')

#plot 1.2
plot(x = d$datetime, 
     y = d$Voltage, 
     type = 'l',
     xlab = '',
     ylab = 'Voltage')

#plot 2.1
plot(x = d$datetime, 
     y = d$Sub_metering_1, 
     type = 'l',
     xlab = '',
     ylab = 'Energy sub meeting',
     col = 'black')
par(new = T)
lines(x = d$datetime, 
     y = d$Sub_metering_2, 
     col = 'red')
par(new = T)
lines(x = d$datetime, 
     y = d$Sub_metering_3, 
     col = 'blue')

legend( 'topright', 
       legend = c('Sub_meeting_1','Sub_meeting_2', 'Sub_meeting_3'),
       col = c('black', 'red', 'blue'),
       lty = 1, 
       cex = 1)
par(new = F)


#plot 2.2
plot(x = d$datetime, 
     y = d$Global_reactive_power, 
     type = 'l',
     xlab = '',
     ylab = 'Global_reactive_power')

```



